---
title: "Week 1 Presentation"
author: "Joleen Vansomphone & Evie Shen"
date: "2024-06-18"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Overview: RStudio

-   Four Quadrants: Console, Script, Environment, Files

a.  Script/Source- Write and store code
b.  Console- Displays the output (contains warnings & errors)
c.  Environment- Where the data appears
d.  Files- Store and open new/existing files (also contains Plots, Packages, Help, Viewer, and Presentation)

-   Major Tools: Run, Broom, Outline

a.  Run - Runs selected line of code (Ctrl + Enter) Use "Source" to run all
b.  Broom - Cleans out console/ environment
c.  Outline - Easy way to navigate documents through notes (begins with \# and ends with dashes)

## Overview: RStudio (For Data Analysis)

-   Install Package: install.packages(“[name of package]”)

a.  Include quotations! (Common error)

-   Common Commands:

1.  View([file name])
2.  dim([file name])
3.  str([file name])
4.  glimpse([file name]) -\> Preferred version of str
5.  head([file name])
6.  tail([file name])
7.  summary([file name])
8.  skim([file name])

## What We Found Most Useful

Useful packages for Data Analysis
(tidyverse, here, janitor, skimr, data.table, nycflights13)
```{r packages}
library(tidyverse)
library(here)
library(janitor)
library(skimr) 
library(data.table) 
library(nycflights13)
```
For Cleaning, Reading, Summarizing

## Useful commands
-   Skim (Great for sorting variables and first impression)
```{r commands}

skim(flights)

```

## Glimpse 
-   Helps especially when data cannot fit 
```{r glimpse}
glimpse(flights)
```
## Summary
Helps with crucial information
```{r summary}
summary(flights)

```

## Overview: Diabetes Research

-   Diabetes occurs when insulin is unable to regulate sugar levels in the bloodstream. 
-   Type 1 Diabetes is where one’s body does not make insulin (autoimmune disorder - caused by an overactive immune system). -  -    Type 2 Diabetes is where one’s body does not react to the insulin produced, called insulin resistance. 

 In 2019, diabetes was the direct cause of 1.5 million deaths and 48% of all deaths due to diabetes occurred before the age of 70 years. It is one of the top 10 global causes of death.
 
## Joleen's Focus: Diabetes Treatments & Epidemiology

## Epidemiology

Epidemiology: - The amount of those with diabetes is projected to increase to 592 million by 2035 (according the the NIH) - Type 1 accounts for under 15% of all cases, and most occur between birth and 14 years old - Nutritional Biomarkers - High-density lipoprotein cholesterol - Protein Sources Insulin-like growth trait on proteins (Hypothesized Vitamin D) after seeing a link - People who are higher risk: BMIs \> 25 T2-those over 35 Family relations to diabetes - Average increase of 3% worldwide per year - Better living conditions T2- More processed diets and sedentary lifestyles T1- Less microorganism exposure, which increases autoimmunity, causing autoreactive cells to destroy insulin-producing cells

## Treatments

Type 1: - Insulin pumps or injections - Blood Sugar Monitoring - Pancreas transplants - Islet Cell transplants - Extracting insulin-producing cells from a donor pancreas, and inserting them into recipient’s liver vein Type 2: - Mostly lifestyle changes - Oral drug treatments - Insulin - Blood Sugar Monitoring Blood Sugar Monitoring: - Type 1 patient often need it more than Type 2, but both commonly use it, especially with insulin therapy - Insulin pumps

## Evie's Focus: Causes & Symptoms

## Causes

Causes: Type 1 Diabetes: Your body does not make enough insulin (autoimmune disorder - caused by an overactive immune system). One’s immune system would mistake cells in their pancreas as invaders, and attack insulin-making cells, destroying their ability to naturally regulate their blood sugar.

Type 2 Diabetes: Your body does not use insulin properly (90% of people with diabetes). This is caused by insulin resistance, when your cells do not respond to insulin. Because the insulin is not being used or responded to, one’s pancreas goes through hyperinsulinemia to make more, but once it is clear that it is ineffective, insulin production is slowed and may be stopped completely.

## Symptoms

Symptoms varies by type of Diabetes (Type 1 Diabetics usually experience more severe symptoms). - Constantly thirsty / Urinating often (increased blood sugar levels cause one’s kidneys to work harder than usual to keep balanced glucose levels and if they cannot keep up, the body will try to flush out excess glucose through urine, causing these symptoms.) - Losing weight (one’s body cannot use the glucose in their blood for energy and must burn fat and muscle instead) - Feeling tired and weak (caused by cells deprived of energy in the form of glucose and dehydration from constant urination) - Feeling irritable or having mood swings (caused by changes in blood sugar levels) - Having blurry vision (Diabetic retinopathy, high blood sugar damages blood vessels in the retina (which swell and leak), causing blurry vision)

## Challenges

-   Figuring out RMarkdown (Looked it up)

-   Updating R 

    ## Ongoing Questions

-   How to load csv files in RMarkdown?

```         
data <- read_csv("Desktop/R Projects Personal/RYouWithMe/Data/sydneybeaches.csv")
data
```

Error: 'Desktop/R Projects Personal/RYouWithMe/Data/sydneybeaches.csv' does not exist in current working directory ('/Users/joleen/Desktop/R Projects Personal/Data Science Internship Presentations').

-   What are the effects of OTC GCM on non-diabetics? 
-   How common are negative side effects of CGMs?

## Week 1

Thank you!
